---
title: "Chile and Peru nutrition optimization walkthrough"
author: "JGM"
date: "5/29/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(message = FALSE)

library (tidyverse)
library (ggradar)

# SAU data
# also just want top ssf spp for each country
sau <- read.csv ("Data/SAU_EEZ_landings.csv")

sau_chl_per <- sau %>%
  filter (area_name %in% c("Chile (mainland)", "Peru"), between (year, 2000, 2015)) %>%
  mutate (country = ifelse (area_name == "Peru", "Peru", "Chile")) %>%
  rename (species = scientific_name)


# nutrient data and rda data
rda_groups <- readRDS("Data/RDAs_5groups.Rds")

fishnutr <- read_csv ("Data/Species_Nutrient_Predictions.csv")
fishnutr_mu <- fishnutr %>%
  select (species, ends_with ("_mu"))

# spp nutrient content
# join to nutrient content
spp_nutr <- fishnutr_mu %>%
  filter (species %in% sau_chl_per$species)  %>%
  pivot_longer (Selenium_mu:Vitamin_A_mu,
                names_to = "nutrient",
                values_to = "amount") %>%
  mutate (nutrient = str_sub(nutrient, end = -4)) %>%
  # join to rda data
  left_join (filter(rda_groups, group == "Child"), by = "nutrient") %>%
  
  # this would be the percentage of your daily requirement you could get from a 100g serving of each species. cap at 100%
  mutate (perc_rda = amount/mean_rda * 100,
          perc_rda = ifelse (perc_rda > 100, 100, perc_rda)) %>%
  ungroup()

spp_nutr_density <- spp_nutr %>% 
  filter (!nutrient %in% c("Protein", "Selenium")) %>%
  select (species, nutrient, perc_rda) %>%
    distinct() %>%
    group_by (species) %>%
    summarise (micronutrient_density = sum (perc_rda))


# full sau nutr data
sau_nutr <- readRDS("Data/SAU_nutr_content_sector_enduse.Rds")


#  nutricast inds fed by sector
proj_inds_fed_sector <- readRDS("Data/Nutricast_by_SAU_sector_inds_fed.Rds")

```

## Project overview

We're exploring how to enhance nutrient provisioning from fisheries. If policymakers in Chile and Peru wanted to harness domestic fisheries to meet their citizens' nutrient needs, what policy interventions should they consider?

* Changing fisheries management to improve yields?
* Keeping nutritious fish in the country instead of exporting it?
* Intervening in supply chains or consumption patterns so the most vulnerable populations are aware of and have access to  the most nutritious species? 


## Data sources


### Past and current landings

I have landings from each country's EEZ from the Sea Around Us project. They show catch by industrial vs. artisanal sector; by foreign vs. domestic vessels; and end use--direct human consumption vs. fishmeal. 

In the following graphs, I'm examining landings from the years 2000-2015. They have reconstructed data from 1950-2018, so the choice of years is arbitrary. The data from 2016-2018 looks funny for all countries, where I don't think they've sorted the fishmeal vs. direct human consumption categories yet. 

* Question for regional teams: Is this set of years appropriate? Were there any big fishery policy changes that would skew the data during these years?

For example, landings over time, by end use:

```{r landings_end_use}
sau_chl_per %>%
  ggplot (aes (x = year, y = tonnes/1000000, fill = end_use_type)) +
  geom_bar(stat = "identity") +
  theme_bw() +
  facet_wrap (~country) +
  labs (x = "", y = "Catch, million tons", fill = "End use")
```


Taking just domestic catch for direct human consumption, the breakdown of industrial and artisanal:

```{r dhc_sector}
sau_chl_per %>%
  filter (end_use_type == "Direct human consumption", fishing_sector %in% c("Industrial", "Artisanal"), fishing_entity %in% c("Chile", "Peru")) %>%
  ggplot (aes (x = year, y = tonnes/1000000, fill = fishing_sector)) +
  geom_bar(stat = "identity") +
  theme_bw() +
  facet_wrap (~country) +
  labs (x = "", y = "Catch, million tons", fill = "Sector")
```

### Exports and trade

Soon, we'll have percentage exports for each country and species from Rashid Sumaila. We hope to also use a new trade database with detailed information about what seafood products are traded. 


### Nutrition

We can then translate that catch into nutritional value. We have fine scale data for finfish based on a predictive model (Christina Hicks's FishNutrients), and coarse data for invertebrates (from the GENuS database; all crustaceans, all cephalopods etc. have the same nutritional values). 

So for example, we can look at the nutrient density of particular species across several micronutrients. In the following graph, the "spokes" represent the percentage of a child's recommended daily nutrient needs they would get from eating that a single 100g serving of each species. 

```{r chl_spp_radar}
# filter nutrient content data
  nutr_radar_plot <- spp_nutr %>%
    filter (species %in% c("Merluccius gayi gayi", "Merluccius australis", "Genypterus maculatus"), !nutrient %in% c("Protein", "Selenium")) %>%
  mutate (comm_name = 
            case_when (species == "Merluccius gayi gayi" ~ "Chilean hake",
                       species == "Merluccius australis" ~ "Austral hake",
                       species == "Genypterus maculatus" ~ "Congrio")) %>%
    select (comm_name, nutrient, perc_rda) %>%
    pivot_wider (
      names_from = nutrient,
      values_from = perc_rda) %>%
    # radar plot can't deal with NA and non-fish don't have selenium
    replace_na (list (Selenium = 0))
  
  
  # set levels so colors stay the same
  nutr_radar_plot$comm_name <- factor (nutr_radar_plot$comm_name, levels = c("Chilean hake", "Austral hake", "Congrio"))
  
   ggradar(nutr_radar_plot,
          grid.min = 0, grid.max = 100, 
          group.point.size = 2,
          group.line.width = 1,
          legend.text.size = 10,
          axis.label.size = 3,
          grid.label.size = 3,
          legend.position = "right") +
    ggtitle ("Child's daily nutrition needs from one serving") +
    theme (plot.title = element_text (size = 14))
  
```
```{r chl_spp_radar_anchov}
# filter nutrient content data
  nutr_radar_plot <- spp_nutr %>%
    filter (species %in% c("Merluccius gayi gayi", "Merluccius australis", "Genypterus maculatus", "Engraulis ringens", "Trachurus murphyi"), !nutrient %in% c("Protein", "Selenium")) %>%
  mutate (comm_name = 
            case_when (species == "Merluccius gayi gayi" ~ "Chilean hake",
                       species == "Merluccius australis" ~ "Austral hake",
                       species == "Genypterus maculatus" ~ "Congrio",
                       species == "Engraulis ringens" ~ "Anchovy",
                       species == "Trachurus murphyi" ~ "Jurel")) %>%
    select (comm_name, nutrient, perc_rda) %>%
    pivot_wider (
      names_from = nutrient,
      values_from = perc_rda) %>%
    # radar plot can't deal with NA and non-fish don't have selenium
    replace_na (list (Selenium = 0))
  
  
  # set levels so colors stay the same
  nutr_radar_plot$comm_name <- factor (nutr_radar_plot$comm_name, levels = c("Chilean hake", "Austral hake", "Congrio", "Anchovy", "Jurel"))
  
   ggradar(nutr_radar_plot,
          grid.min = 0, grid.max = 100, 
          group.point.size = 2,
          group.line.width = 1,
          legend.text.size = 10,
          axis.label.size = 3,
          grid.label.size = 3,
          legend.position = "right") +
    ggtitle ("Child's daily nutrition needs from one serving") +
    theme (plot.title = element_text (size = 14))
  
```

We can summarize the overall "micronutrient density" of a species by taking the sum of the values across the five spokes, where the maximum micronutrient density level is 500 and the minimum is zero. Of the species caught in Chile and Peru, the highest value is 287, the lowest is 37, and the mean is 96. For the species above, Chilean hake has a micronutrient density of 44; Austral hake, 47; congrio, 71; jurel, 152; and anchovy, 170.

I'll use these values later when trying to prioritize specific species. 
```{r sau_sector_density, include = FALSE}
g <- sau_chl_per %>%
  group_by (country, species, end_use_type) %>%
  summarise (tonnes_tot = sum (tonnes)) %>%
  left_join (spp_nutr_density, by = "species") %>%
  filter (!is.na (micronutrient_density)) %>%
  group_by (country, end_use_type) %>%
  summarise (density_wt = weighted.mean (micronutrient_density, weights = tonnes_tot))


h <- sau_chl_per %>%
  group_by (country, species, fishing_sector) %>%
  summarise (tonnes_tot = sum (tonnes)) %>%
  left_join (spp_nutr_density, by = "species") %>%
  filter (!is.na (micronutrient_density)) %>%
  group_by (country, fishing_sector) %>%
  summarise (density_wt = weighted.mean (micronutrient_density, weights = tonnes_tot))

k <- sau_chl_per %>%
  group_by (country, species, end_use_type) %>%
  summarise (tonnes_tot = sum (tonnes)) %>%
  left_join (spp_nutr_density, by = "species") %>%
  filter (!is.na (micronutrient_density)) %>%
  mutate ()

```

### Country nutritional needs

We have estimates from Nutricast of each country's nutrition deficits based on its food supply, population size, and recommended daily intake values of nutrients. We've also looked for specific health policy priorities in each country. From Nutricast, calcium and Vitamin A are priorities for Chile; additional studies showed that Iron and Zinc are also priorities. For Peru, Nutricast showed that calcium is a priority, and additional studies revealed deficiencies in Vitamin A and Iron. 



We can convert current landings into their nutritional value to show how current catch could contribute to nutrient needs. For example, how many children's nutritional needs could be met each year from current landings (domestic, direct human consumption).

```{r sau_landings_nutr_content}

t <- sau_nutr %>%
  filter (country %in% c("Chile", "Peru"), end_use_type == "Direct human consumption", fishing_entity %in% c("Chile", "Peru"), fishing_sector %in% c("Industrial", "Artisanal")) %>%
  group_by (country, fishing_sector, nutrient, year) %>%
  summarise (annual_nutr_servings = sum(nutr_servings, na.rm = TRUE)) %>%
  group_by (country, fishing_sector, nutrient) %>%
  summarise (avg_nutr_servings = mean (annual_nutr_servings)) %>%
  left_join (filter (rda_groups, group == "Child"), by = "nutrient") %>%
  mutate (needs_met = avg_nutr_servings / mean_rda)
  

t %>%
  filter (!nutrient %in% c("Protein", "Selenium")) %>%
  ggplot (aes (x = nutrient, y = needs_met/1000000, fill = fishing_sector)) +
  geom_bar (stat = "identity", position = "dodge") +
  facet_wrap (~country) +
  theme_bw() +
  labs (x = "Nutrient", y = "Children fed, millions", fill = "Sector")

```

### Future landings under climate change and management scenarios

Finally, what might happen to that nutrition provisioning under climate change? Can we improve nutrition provisioning with adaptive fisheries management? We can use EDF/UCSB's climate upside model and translate projected changes in yield into nutrients. 

Projected number of children fed each year by mid-century under the climate scenario RCP 6.0, with SSF catch
```{r nutricast}
proj_inds_fed_sector %>%
  filter (group == "Child", country %in% c("Chile", "Peru"), rcp == "RCP60", scenario %in% c("No Adaptation", "Full Adaptation"), period == "2050-2060", !nutrient %in% c("Protein", "Selenium")) %>%
  ggplot (aes (x = nutrient, y = inds_fed_ssf/1000000, fill = scenario)) +
  geom_bar (stat = "identity", position = "dodge") +
  theme_bw() +
  facet_wrap (~country) +
  labs (x = "Nutrient", y = "Children fed, millions", fill = "Management\n scenario")

```



## Prioritizing species

Since it's overwhelming to look at overall landings made up of dozens of species, we're hoping to focus on a subset (< 10) of priority species for each country. 

We could choose species based on:
* Which species show the biggest future gains from adaptive management
* Which species provide needed nutrients
* Important SSF species by volume or value
* Other criteria?

