---
title: "SSF nutrient contribution explore"
author: "JGM"
date: "4/1/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

library (tidyverse)
library (RColorBrewer)


# datasets ----
# Sau data 
sau_ds <- readRDS("Data/SAU_countries_dirhumcons_2000_2015.Rds")

# sau species nutrient data
sau_spp_nutr <- readRDS("Data/SAU_spp_nutrient_data.Rds") 

# sau catch data with nutrients
sau_nutr <- readRDS("Data/SAU_nutr.Rds")%>%
  ungroup() %>%
  filter (year < 2016)

# rdas
rda_groups <- readRDS("Data/RDAs_5groups.Rds")

# population data ----

# downloaded from https://population.un.org/wpp/Download/Standard/CSV/ on 3/17/2022
# Population by 5-year age groups.
# PopMale: Male population in the age group (thousands)
# PopFemale: Female population in the age group (thousands)
# PopTotal: Total population in the age group (thousands)

pop <- read_csv("Data/WPP2019_PopulationByAgeSex_Medium.csv")

# separate into broad adult male/female and child categories to  match RDAs
pop_2020 <- pop %>%
  filter (Location %in% c("Sierra Leone", "Indonesia", "Peru", "Chile", "Malawi"), Time ==2020) %>%
  mutate (group = 
            case_when (
              AgeGrp %in% c("0-4", "5-9") ~ "Child",
              !AgeGrp %in% c("0-4", "5-9") ~ "Adult")
  ) %>%
  group_by (Location) %>%
  summarize (Pop_Males =  1000 * sum (PopMale[group == "Adult"]),
             Pop_Females = 1000 * sum(PopFemale[group == "Adult"]),
             Pop_Child = 1000 * sum(PopTotal[group == "Child"])) %>%
  # also have total population that would eat fish--sum of males, females, and children maybe do this later?
  #Pop_Total = sum (Pop_Males, Pop_Females, Pop_Child)) %>%
  pivot_longer (cols = starts_with("Pop"),
                names_prefix = "Pop_", 
                names_to = "group", 
                values_to = "population") %>%
  rename (country = Location)

# climate vulnerability and fishing vulnerability data
# can prob get clim vuln from fishbase, steal from maire for now
# maire_vuln <- read_csv ("../NutrientGlobalFisheries_Maire/data/species_data.csv") %>%
#   select (species, Fishing_V, Climate_V.index) %>%
#   rename (fishing_vuln = Fishing_V, climate_vuln = Climate_V.index)
# 


# functions ----
# Function to calculate mt of nutrient from mt of edible meat ----
# units: mg, ug=mcg 
# meat_mt <- 29.88111; nutr_dens <- 35.5; nutr_dens_units <- "mg"
# from nutrient_endowment --> shiny --> v3 --> :Page3_Fig2c_reforms_prop_demand, line 84

# adding--divide by 100 in here. I think Chris did that in his head but making me doubt everything
calc_nutr_supply_mt <- function(meat_mt, nutr_dens, nutr_dens_units){
  
  # Convert meat to grams
  # "Mg" is metric tons
  meat_g <- measurements::conv_unit(meat_mt, "Mg", "g")
  
  # Calculate amount of nutrient in density units. divide by 100 because density units are per 100g
  nutrient_q <- meat_g *  nutr_dens / 100
  
  # Calculate amount of nutrient in metric tons
  nutrient_mt <- measurements::conv_unit(nutrient_q, nutr_dens_units, "Mg")
  
  # Return
  return(nutrient_mt)
  
}
```

I downloaded Sea Around Us data for our focal countries. I consolidated the three Indonesia regions (central, Indian ocean, east), and only used mainland Chile, not Easter Island. 

I only used "landings", not "discards." I am only looking at fish for "direct human consumption", not fishmeal/fish oil. The industrial fishing sector may include foreign catch. This is fairly small, but we could look at the nutrition loss to foreign catch. 

I'm looking at recent data, from 2001-2015. In the most recent years, 2016-2018, they haven't split out fishmeal and oil for any country. 

## Can we draw a clear line between industrial and artisanal species?
```{r sector_spp}
sector_spp <- sau_ds %>%
  ungroup() %>%
  rename (species = scientific_name) %>%
  #filter (country =="Chile") %>%
  group_by (country, species) %>%
  # define as ssf when > 75% of catch is in SSF 
  mutate (spp_sector = case_when (
    sum (tonnes_tot[fishing_sector == "Industrial"])/sum (tonnes_tot) > 0.75 ~ "Industrial",
    sum (tonnes_tot[fishing_sector == "Artisanal"])/sum (tonnes_tot) > 0.75 ~ "SSF",
    TRUE ~ "Mixed")
    ) %>%
  select (country, species, spp_sector) %>%
  distinct()

#table (chl_sector_spp$spp_sector)
sector_spp %>%
  group_by (country, spp_sector) %>%
  summarise (n_spp = n())

saveRDS(sector_spp, file = "Data/SAU_spp_by_sector_75cutoff.Rds")

t90 <- sau_ds %>%
  ungroup() %>%
  rename (species = scientific_name) %>%
  #filter (country =="Chile") %>%
  group_by (country, species) %>%
  # define as ssf when > 75% of catch is in SSF 
  mutate (spp_sector = case_when (
    sum (tonnes_tot[fishing_sector == "Industrial"])/sum (tonnes_tot) > 0.9 ~ "Industrial",
    sum (tonnes_tot[fishing_sector == "Artisanal"])/sum (tonnes_tot) > 0.9 ~ "SSF",
    TRUE ~ "Mixed")
    ) %>%
  select (country, species, spp_sector) %>%
  distinct()

table (t90$spp_sector)

t90 %>%
  group_by (country, spp_sector) %>%
  summarise (n_spp = n())

```
With a 75% cutoff
Chile: 30 industrial, 85 ssf, 4 mixed. 
Indo: 114 indus, 47 mixed, 16 SSF
Peru: 35 industrial, 38 ssf, 6 mixed
SL: 142 industrial, 15 mixed, 45 ssf

With a 90% cutoff:
Chile- 29 industrial, 82 SSF, 8 mixed
Indo- 96 industrial, 69 mixed, 12 SSF --> not as good
Peru - 32 industrial, 31 SSF, 16 mixed
SL: 142 industrial, 21 mixed, 39 SSF

maybe use 75


## Overall micronutrient density of species by sector

We define overall micronutrient density of a species from Maire et al. (2021): the micronutrient density for a specific nutrient would be the percentage of a child's recommended dietary allowance of that nutrient in a 100g serving of that species, e.g. if a 100g serving of cod provides 50% of the Iron RDA for children, its iron density is 50. The overall micronutrient density is the sum for all nutrients we're examining--calcium, vitamin A, selenium, iron, zinc, and omega 3s. The maximum value is 600 and the minimum value is zero. 

```{r micro_dens}
### plot distribution of micronutrient density by sector
spp_micronutr_density <- spp_nutr %>%
  filter (group == "Child", nutrient != "Protein") %>%
  distinct() %>%
  group_by (species) %>% 
  summarise (micronutrient_density = sum (perc_rda, na.rm = TRUE))

sau_ds %>%
  rename (species = scientific_name) %>%
  left_join (spp_micronutr_density, by = "species") %>%
  ggplot () +
  geom_density (aes (x = micronutrient_density, col = fishing_sector), alpha = 0.5) +
  facet_wrap (~country) +
  theme_bw() +
  ggtitle ("Distribution of micronutrient density of species caught by each fishing sector")


# can I make a density graph or a line graph where the x axis is micronutrient density and the y axis is tonnes caught and the fill is sector.

# https://stackoverflow.com/questions/20342494/density-of-each-group-of-weighted-geom-density-sum-to-one
sau_ds %>%
  rename (species = scientific_name) %>%
  left_join (spp_micronutr_density, by = "species") %>%
  group_by (fishing_sector) %>%
  mutate (tonnes_sector = sum(tonnes_tot)) %>%
  ggplot () +
  geom_density (aes (x = micronutrient_density,  col = fishing_sector, group = fishing_sector, weight = tonnes_tot/tonnes_sector), alpha = 0.5) +
  facet_wrap (~country) +
  theme_bw() +
  ggtitle ("Distribution of micronutrient density of species caught by each fishing sector, weighted by catch")
```


compare micronutrient density, climate vulnerability, fishing vulnerability

```{r vulnerability_micro_dens}

spp_profiles_vuln <- spp_micronutr_density %>%
left_join (maire_vuln, by = "species")

ggplot (spp_profiles_vuln) +
  geom_point (aes (x = fishing_vuln, y = micronutrient_density, col = climate_vuln)) +
  theme_bw() +
  scale_color_distiller(palette = "RdYlGn")

```

relate this to ssf vs. artisanal?
with 75% cutoff, 30 industrial species, 4 mixed, 85 SSF. 
```{r chl_spp_vuln_profile_sector}



spp_profiles_vuln %>%
  left_join (t, by = "species") %>%
  filter (!is.na (spp_sector)) %>%
  ggplot () +
  geom_point (aes (x = fishing_vuln, y = micronutrient_density, col = climate_vuln, shape = spp_sector)) +
  theme_bw() +
  scale_color_distiller(palette = "RdYlGn") +
  ggtitle ("Chile species")


          
spp_profiles_vuln %>%
  left_join (t, by = "species") %>%
  filter (!is.na (spp_sector)) %>%
  ggplot () +
  geom_boxplot (aes (x = spp_sector, y = micronutrient_density)) +
  theme_bw() +
  ggtitle ("Chile species")

spp_profiles_vuln %>%
  left_join (t, by = "species") %>%
  filter (!is.na (spp_sector)) %>%
  ggplot () +
  geom_boxplot (aes (x = spp_sector, y = climate_vuln)) +
  theme_bw() +
  ggtitle ("Chile species")

spp_profiles_vuln %>%
  left_join (t, by = "species") %>%
  filter (!is.na (spp_sector)) %>%
  ggplot () +
  geom_boxplot (aes (x = spp_sector, y = fishing_vuln)) +
  theme_bw() +
  ggtitle ("Chile species")
```
what about 90% as cutoff?
29 industrial, 8 mixed, 82 ssf. may as well do 90. 
```{r chl_spp_vuln_profile_sector_90cutoff}



# quick check since anchovy does a big jump to industrial in 2015. 
sau_ds %>%
  filter (scientific_name %in% t90$species[which (t90$spp_sector == "Mixed")]) %>%
  ggplot (aes (x = year, y = tonnes_tot, fill = fishing_sector)) +
  geom_bar(stat = "identity") + 
  theme_bw() +
  facet_wrap (~scientific_name)
```
d gigas is pretty evenly mixed. anchovy for dhc is almost all artisanal, than does a big jump to industrial in 2014. 
``` {r plot_chile_sector_attributes}
spp_profiles_vuln %>%
  left_join (t90, by = "species") %>%
  filter (!is.na (spp_sector)) %>%
  ggplot () +
  geom_point (aes (x = fishing_vuln, y = micronutrient_density, col = climate_vuln, shape = spp_sector)) +
  theme_bw() +
  scale_color_distiller(palette = "RdYlGn") +
  ggtitle ("Chile species")


#           
spp_profiles_vuln %>%
  left_join (t90, by = "species") %>%
  filter (!is.na (spp_sector)) %>%
  ggplot () +
  geom_boxplot (aes (x = spp_sector, y = micronutrient_density)) +
  theme_bw() +
  ggtitle ("Micronutrient density of species by sector, Chile")

spp_profiles_vuln %>%
  left_join (t90, by = "species") %>%
  filter (!is.na (spp_sector)) %>%
  ggplot () +
  geom_boxplot (aes (x = spp_sector, y = climate_vuln)) +
  theme_bw() +
  ggtitle ("Climate vulnerability of species by sector, Chile")

spp_profiles_vuln %>%
  left_join (t90, by = "species") %>%
  filter (!is.na (spp_sector)) %>%
  ggplot () +
  geom_boxplot (aes (x = spp_sector, y = fishing_vuln)) +
  theme_bw() +
  ggtitle ("Fishing vulnerability of species by sector, Chile")
```
Mixed are most nutritious. SSF has slightly higher mean than industrial but bigger spread. SSF more climate vulnerable, industrial most fishing vulnerable. 

```{r most_nutritious_chile}

spp_profiles_vuln %>%
  left_join (t90, by = "species") %>%
  filter (!is.na (spp_sector)) %>%
  slice_max (micronutrient_density, prop = 0.25) %>%
  ggplot () +
  geom_boxplot (aes (x = spp_sector, y = climate_vuln)) +
  theme_bw() +
  ggtitle ("Climate vulnerability of most nutritious species by sector, Chile")

spp_profiles_vuln %>%
  left_join (t90, by = "species") %>%
  filter (!is.na (spp_sector)) %>%
  slice_max (micronutrient_density, prop = 0.25) %>%
  ggplot () +
  geom_boxplot (aes (x = spp_sector, y = fishing_vuln)) +
  theme_bw() +
  ggtitle ("Fishing vulnerability of most nutritiousspecies by sector, Chile")

```


what about the climate/fishing vulnerability of the most nutritious spp--similar patterns but more extreme. 

## break this out by individual nutrients
```{r chl_spp_sector_by_nutrient}

t90 %>%
  left_join (spp_nutr, by = "species") %>%
  filter (group == "Child", !nutrient %in% c("Protein")) %>%
  ggplot () +
  geom_boxplot (aes (x = spp_sector, y = perc_rda)) +
  facet_wrap (~nutrient) +
  theme_bw()
  

```
anchovy is in mixed, that might be driving omega 3s? slightly higher vit A for ssf?



### Current SAU nutrient yield and contribution to population needs

not based on species, just break out SAU sectors. Use the 2020 population estimate, but take average catch 2000-2015
```{r catch_nutr_contribution}


# just want population level rda for now 
pop_rda_needs <- pop_2020 %>%
  left_join (rda_groups, by = "group") %>%
  mutate (rda_needs = mean_rda * population) %>%
  group_by (country, nutrient) %>%
  summarise (
    tot_rda_needs = sum (rda_needs, na.rm = TRUE))

# take average nutrient servings by nutrient and sector

sau_chl_test <- sau_nutr %>%
  filter (country == "Chile", group == "Child", species %in% c("Merluccius gayi gayi", "Engraulis ringens")) %>%
  select (year, species, fishing_sector, tonnes_tot, nutrient, amount, group, perc_rda, meat_mt, nutr_mt, meat_servings, nutr_servings)

z <- sau_chl_test %>% 
group_by (species, fishing_sector, nutrient) %>%
  summarise (mean_tonnes = mean (tonnes_tot, na.rm = TRUE),
          mean_servings = mean (nutr_servings, na.rm = TRUE))

z2 <- sau_chl_test %>% 
group_by (fishing_sector, nutrient) %>%
  summarise (mean_tonnes = mean (tonnes_tot, na.rm = TRUE),
          mean_servings = mean (nutr_servings, na.rm = TRUE))

z_rda <- z %>%
 left_join (filter(pop_rda_needs, country == "Chile"), by = c("nutrient")) %>%
  mutate (#yield_per_cap = nutr_servings / population,
          
          # proportion of population rda met for each group
          rda_met = mean_servings/tot_rda_needs)

z2_rda <- z2 %>%
left_join (filter(pop_rda_needs, country == "Chile"), by = c("nutrient")) %>%
  mutate (#yield_per_cap = nutr_servings / population,
          
          # proportion of population rda met for each group
          rda_met = mean_servings/tot_rda_needs)
# this is much smaller, this is the mean of all the species instead of the sum. should use the z approach. 

ggplot (z1_rda, aes (y = rda_met, x = nutrient, fill = fishing_sector)) +
  geom_bar (stat = "identity", position = "dodge") +
  facet_wrap (~country, nrow = 4, scales = "free") +
  theme_bw() +
  labs (x = "", fill = "", y = "") +
  facet_wrap (~country, scales = "free") +
  labs (y = "Proportion of population RDAs met") +
  ggtitle ("Proportion of population RDAs met by current catch\nSAU data, average catch 2000-2015") +
  theme (axis.text.x = element_text (angle = 60, hjust = 0.9))

# check if reasonable
ggplot (sau_chl_test, aes (x = year, y = tonnes_tot, col = species)) + geom_point() + geom_line() + theme_bw() + facet_wrap (~fishing_sector)

rdas_pop_level_sau <- sau_nutr %>%
  group_by (country, fishing_sector, species, nutrient) %>%
  mutate (mean_tonnes = mean (tonnes_tot, na.rm = TRUE),
          mean_servings = mean (nutr_servings, na.rm = TRUE)) %>%
  select (country, species, major_group, fishing_sector, nutrient, mean_servings) %>%
  distinct() %>%
 left_join (pop_rda_needs, by = c("country", "nutrient")) %>%
  mutate (#yield_per_cap = nutr_servings / population,
          
          # proportion of population rda met for each group
          rda_met = mean_servings/tot_rda_needs)
 

rdas_pop_level_sau %>%
  filter (country %in% c("Chile", "Peru"), 
          fishing_sector %in% c ("Artisanal", "Industrial")) %>%
  ggplot (aes (y = rda_met, x = nutrient, fill = fishing_sector)) +
  geom_bar (stat = "identity", position = "dodge") +
  facet_wrap (~country, nrow = 4, scales = "free") +
  theme_bw() +
  labs (x = "", fill = "", y = "") +
  facet_wrap (~country, scales = "free") +
  labs (y = "Proportion of population RDAs met") +
  ggtitle ("Proportion of population RDAs met by current catch\nSAU data, average catch 2000-2015") +
  theme (axis.text.x = element_text (angle = 60, hjust = 0.9))





```



there's a weird thing in the Peru data where from 2016-2018 there's a huge jump in industrial-caught anchovy going to direct human consumption. all industrial catch in those years is direct human consumption, whereas usually it's mostly fishmeal. 


chile data still weird, huge jump in ahcnovy in 2014 2015 too

```{r peru_industrial_anchovy}

sau_ds %>%
  filter (country == "Peru") %>%
  mutate (anchov = ifelse (scientific_name == "Engraulis ringens", "Anchovy", "Other")) %>%
  ggplot (aes (x = year, y = tonnes_tot, fill = anchov)) +
  geom_bar (stat = "identity") +
  facet_wrap (~fishing_sector, scales = "free")

```


