---
title: "Nutrition optimization modeling--plots for review"
author: "JGM"
date: "3/18/2022"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
library (tidyverse)

# for nested facet_wrap for plotting
#devtools::install_github("teunbrand/ggh4x")
library (ggh4x)


# data tables

# Micronutrient deficiencies
mic_def <- read_csv ("Data/mic_def_compare_beal_golden.csv")

# nutrient content of each species with rda met
ds_spp_nutr_content <- readRDS("Data/ds_spp_nutr_content_FishNutrientsGENuS_RDA_groups.Rds")

# nutrient yield of baseline catch (2012-2020, BAU) in mt and servings
ds_catch_nutr_yield_baseline <- readRDS("Data/ds_catch_nutr_yield_baseline.Rds")

# rdas met, baseline catch (2012-2020, BAU). at species level
ds_rdas_met_baseline <- readRDS("Data/ds_baseline_RDAs_met.Rds")

# rdas met, future proections. at species level
ds_rdas_met_projected <- readRDS ("Data/ds_projected_RDAs_met.Rds")

#demand_met_allscen <- readRDS("Data/nutr_demand_met_allscen.Rds")


# functions for plotting
source ("Scripts/Plot_spp_nutr_density.R")
# source ("Scripts/Plot_nutr_demand_met.R")
# source ("Scripts/Plot_n_inds_reqs_met.R")


# plot RDAs met under management scenarios
plot_rdas_met_scen <- function (country_name, imperfect = TRUE, anchovy = TRUE) {
  
  
  
  # Do we want the imperfect scenarios shown?
  if (imperfect == TRUE) {
    country_rdas_met <- ds_rdas_met_projected %>%
      filter (country == country_name) %>%
      group_by (rcp, scenario, period, nutrient, species) %>%
  mutate (tot_pop = sum(population),
          tot_rda_needs = sum(rda_needs),
          tot_nutr_servings = first (nutr_servings),
          tot_rda_met = nutr_servings / tot_rda_needs) %>%
  # get rid of groups, otherwise triple counting
  select (country, rcp, scenario, period, species, nutrient, tot_rda_met) %>%
  distinct() 
    
    country_rdas_met$scenario <- factor (country_rdas_met$scenario, levels = c("No Adaptation", "Productivity Only", "Range Shift Only", "Imperfect Productivity Only", "Imperfect Full Adaptation", "Full Adaptation"))
  
    
  } else {
    country_rdas_met <- ds_rdas_met_projected %>%
      filter (country == country_name,
              !grepl("Imperfect", scenario)) %>%
      group_by (rcp, scenario, period, nutrient, species) %>%
  mutate (tot_pop = sum(population),
          tot_rda_needs = sum(rda_needs),
          tot_nutr_servings = first (nutr_servings),
          tot_rda_met = nutr_servings / tot_rda_needs) %>%
  # get rid of groups, otherwise triple counting
  select (country, rcp, scenario, period, species, nutrient, tot_rda_met) %>%
  distinct() 
    
    country_rdas_met$scenario <- factor (country_rdas_met$scenario, levels = c("No Adaptation", "Productivity Only", "Range Shift Only", "Full Adaptation"))
  }
  
  # include anchovy?
  if (anchovy != TRUE) {
    country_rdas_met <- country_rdas_met %>%
      filter (species != "Engraulis ringens")
    
    title = paste0("Percent of population RDAs met under future management scenarios,\n ", country_name, ", anchovy removed")
  } else {
    
    title = paste0("Percent of population RDAs met under future management scenarios,\n", country_name)
    
  }
  
  
  # set period levels
  country_rdas_met$period <- factor(country_rdas_met$period, levels = c("2025-2035","2050-2060", "2090-2100"))
  
  # plot
  country_rdas_met %>%
  ggplot (aes (y = tot_rda_met * 100, x = period, fill = scenario)) +
  geom_bar (stat = "identity", position = "dodge") +
  facet_grid (nutrient ~ rcp, scales = "free") +
  theme_bw() +
  theme (axis.text = element_text(angle = 0.6)) +
  labs(y="Percent of RDAs met from fisheries reforms", x="", fill = "Management\nscenario") +
  ggtitle (title)
  
  
}


# vector of country names
demonstration_countries <- c("Chile", "Peru", "Indonesia", "Sierra Leone")
```

## Micronutrient deficiencies in case study countries

See "Micronutrient deficiencies" table in google doc for prevalence of inadequate intake values from two sources, Nutricast and Beal et al., (2017). I have deficiencies for Calcium, Vitamin A, Zinc, and Iron (these are also what Hicks et al looked at). For all our countries, calcium has the highest prevalence of deficiency, followed by Vitamin A for every country but Peru (Zinc, then Vitamin A). However, since there are only four nutrients to look at, and because these data don't necessarily reflect what we've found in our country-specific research (e.g. concerns about anemia in Peru and iron and zinc deficiencies in Chile), *we might as well look at all of them.* In fact, I'm not sure that it's worth leaving out Selenium and Omega-3s either, which probably also represent deficiencies but just have fewer intake or deficiency studies. 

Nutricast used nutrient supply from the GENuS database, Estimated Average Requirements (EARs), which represent the nutrients needed to meet the needs of half the population, and sex/age specific estimated population size  to calculate micronutrient deficiencies. They used an optimization function to expand the EARs to reach 95% of the nutritional needs of the population. I have access to this code so can get age/sex specific deficiency values. However, this calculation calls for much higher values than the Recommended Dietary Allowance (RDA) values, which are supposed to meet the needs of 97-98% of the population (Hicks etc. use the RDA values, which I think are the same as the RNI values that Illuminating Hidden Harvest uses).

Beal et al. (2017) calculate a Prevalence of Inadequate Micronutrient Intake Index (PIMII) using I believe the same method as above (nutricast probably modeled after them), based on EAR cutpoints, but gets different values than Nutricast. They also provide estimates accounting for fortification in diets. These data are from 2011 and do not include selenium or Omega 3s. Hicks and associated authors/papers (Maire, Robinson) use these values, which are available as overall index and by individual nutrients from Beal's supplemental data. I therefore only have population-level values, not age/sex specific. 


For further reference, Chris Golden's AFCD/Blue Foods Assessment paper used the proprietary Global Nutrient Database (based on FAO Supply and Utilization Accounts) for national nutrient availability, and GENuS for intake/consumption. They used a habitual intake statistical program (SPADE) to calculate consumption of Omega 3s and vitamin B12, which are not in GENuS. They used SEVs (summary exposure values) as a risk-weighted prevalence of inadequate micronutrient intake. I'm not sure yet what the benefit is of this method over the others. 


## Nutrient density of major species by catch volume

These cobweb plots tell us about the nutrient content of the most-caught species in each country. These are slightly different from a version I showed earlier, and I think more relevant. Now the y axis is the % recommended dietary allowance (RDA) for a child age 6mos to 8 years old from a 100g serving of that species. 100% means that eating a single serving of that species would satisfy a child's daily needs for that nutrient. 



```{r spider_plots, echo = FALSE}

radar_plots <- lapply (demonstration_countries, plot_spp_nutr_radar, n_spp = 5)
#grid.arrange (radar_plots[[1]], radar_plots[[2]],radar_plots[[3]],radar_plots[[4]], ncol =1)

radar_plots[[1]]

```
```{r, echo = FALSE}
# surely a better way of doing this??
radar_plots[[2]]
```
```{r, echo = FALSE}
radar_plots[[3]]
```

Note: Acetes japonicus and P. trituberculatus are both crustaceans and thus have identical nutrient profiles from GENuS data. Last graph had a different fish, was 5th most caught with just 2012 while P. trituberculatus was 6th. Can look into further..

```{r, echo = FALSE}
radar_plots[[4]]
```

### How much are the most nutritious species contributing to catch?

```{r catch_vs_density, echo = FALSE}

# truncate name function in plot_spp_nutr_density.R
catch_micronutr_density <- ds_spp_nutr_content %>%
  filter (group == "Child", !is.na (perc_rda)) %>%
  group_by (country, species) %>%
  mutate (micronutrient_density = sum(perc_rda),
          label = ifelse (micronutrient_density > 450 | catch_mt > 500000, 1, 0), 
          name_sm = truncate_name(species)) %>%
  select (country, species, catch_mt, major_group, micronutrient_density, label, name_sm) %>%
  distinct() %>%
  ungroup()

ggplot (catch_micronutr_density, aes (x = log(catch_mt), y = micronutrient_density)) +
  geom_point (aes(col = major_group)) +
  geom_text (data = filter (catch_micronutr_density, label == 1), aes (label = name_sm)) +
  theme_bw() +
  facet_wrap (~country, scales = "free") +
  labs (y = "Micronutrient density", x = "log (Catch, mt)", col = "") +
  ggtitle ("Micronutrient density vs current catch volume")

```

### What is the current nutrient yield under BAU?


Note different scales in y axes, but the values are all in metric tons. I'm separating the nutrients out into the relevant units for human health so they're more visible--protein and omega 3s are measured in grams per serving; CA, Iron, and Zinc in mg; Selenium and Vitamin A in ug. 


```{r nutr_yield_BAU, echo = FALSE, warning = FALSE, fig.width=10,fig.height=14}

# https://stackoverflow.com/questions/62652308/combine-multiple-facet-strips-across-columns-in-ggplot2-facet-wrap

ds_catch_nutr_yield_baseline %>%
  ggplot (aes (x = nutrient, y = nutr_mt, fill = major_group)) +
  facet_nested_wrap (country~ dens_units, ncol = 3, scales = "free") +
  geom_bar (stat = "identity") +
  theme_bw() +
  labs (y = "Yield, mt", fill = "", x = "") +
  ggtitle ("Current nutrient yields under BAU")


```

We can display this instead as the proportion of the population's RDAs (recommended dietary allowances) could be met each day with current catch. The total population RDA accounts for age and sex: I took mean RDA values for children (age 6 mo to 8 years), men, and women, multiplied them by the population numbers for those groups, and took the sum of those three values.



```{r rdas_met_total, echo = FALSE}

rdas_pop_level <- ds_rdas_met_baseline %>%
  filter (!is.na (nutr_servings)) %>%
  group_by (country, nutrient, species) %>%
  mutate (tot_pop = sum(population),
             tot_rda_needs = sum(rda_needs),
             tot_nutr_servings = first (nutr_servings),
             tot_rda_met = nutr_servings / tot_rda_needs) %>%
  # get rid of groups, otherwise triple counting
  select (country, species, major_group, nutrient, tot_rda_met) %>%
  distinct() 

rdas_pop_level %>%
  ggplot (aes (y = tot_rda_met, x = nutrient, fill = major_group)) +
  geom_bar (stat = "identity") +
  facet_wrap (~country, nrow = 4, scales = "free") +
  theme_bw() +
  labs (x = "", fill = "", y = "") +
  facet_wrap (~country, scales = "free") +
  labs (y = "Proportion of population RDAs met") +
  ggtitle ("Proportion of population RDAs met by current catch, BAU") +
  theme (axis.text.x = element_text (angle = 60, hjust = 0.9))

```


For Peru and Chile, this is mostly anchovy: 

```{r rdas_met_noanchov, echo = FALSE}

rdas_pop_level %>%
  mutate(
          anchovy = ifelse (species == "Engraulis ringens", "Anchovy", "Other")
          ) %>%
 ggplot (aes (y = tot_rda_met, x = nutrient, fill = anchovy)) +
  geom_bar (stat = "identity") +
  facet_wrap (~country, nrow = 4, scales = "free") +
  theme_bw() +
  facet_wrap (~country, scales = "free") +
  labs (y = "Proportion of population RDAs met", fill = "", x = "") +
  ggtitle ("Proportion of population RDAs met by current catch, BAU")+
  theme (axis.text.x = element_text (angle = 60, hjust = 0.9))


rdas_pop_level %>%
  filter (species != "Engraulis ringens") %>% 
 ggplot (aes (y = tot_rda_met, x = nutrient, fill = major_group)) +
  geom_bar (stat = "identity") +
  facet_wrap (~country, nrow = 4, scales = "free") +
  theme_bw() +
  labs (x = "", fill = "", y = "") +
  facet_wrap (~country, scales = "free") +
  labs (y = "Proportion of population RDAs met") +
  ggtitle ("Proportion of population RDAs met by current catch, BAU \nAnchovy removed")+
  theme (axis.text.x = element_text (angle = 60, hjust = 0.9))

```

We can also break out the population into groups--what it would look like if we fed all the seafood to children, for example. Unfortunately don't have population breakdowns for pregnant and lactating women, but we could look at the elderly. These proportions are larger because it represents what would happen if ALL the current catch were provided to these groups; e.g. only children were eating fish. 

```{r rdas_by_pop_group, echo = FALSE}

ds_rdas_met_baseline %>%
  filter (!is.na(rda_met), species != "Engraulis ringens") %>%
  ggplot (aes (y = rda_met, x = nutrient, fill = group)) +
  geom_bar (stat = "identity", position = "dodge") +
  facet_wrap (~country, nrow = 4, scales = "free") +
  #geom_text (aes (label = round(rda_met, 1), y = rda_met* 1.05), position = position_dodge(width = 1)) +
  theme_bw() +
  labs (x = "", fill = "", y = "Proportion of population group RDAs met") +
  ggtitle ("RDAs met for each population group, Anchovy removed")

```

## What is the nutritional upside of different management scenarios?

In the previous set of figures, I showed percent of nutrient demand met under the various scenarios using the Nutricast method. Here instead I'm using RDAs. We get somewhat bigger numbers because the RDA values are smaller than the Nutricast optimization method, but not a huge difference. 

```{r SL_future_rdas, echo = FALSE, fig.height=12}

plot_rdas_met_scen("Sierra Leone")

```

```{r In_future_rdas, echo = FALSE, fig.height=12}

plot_rdas_met_scen("Indonesia", imperfect = FALSE)

```

```{r chile_future_rdas, echo = FALSE, fig.height=12}

plot_rdas_met_scen("Chile", imperfect = FALSE)

```

```{r chile_future_rdas_noanchov, echo = FALSE, fig.height=12}

plot_rdas_met_scen("Chile", imperfect = FALSE, anchovy = FALSE)

```

```{r Peru_future_rdas, echo = FALSE, fig.height=12}

plot_rdas_met_scen("Peru", imperfect = FALSE, anchovy = FALSE)

```