---
title: "Nutrition optimization modeling--plots for review"
author: "JGM"
date: "3/18/2022"
output: html_document
---

```{r setup, include=FALSE}
library (tidyverse)


# data tables
ds_spp_nutr_content <- readRDS("Data/ds_spp_nutr_content_FishNutrientsGENuS_RDA_groups.Rds")
#demand_met_allscen <- readRDS("Data/nutr_demand_met_allscen.Rds")


# functions for plotting
source ("Scripts/Plot_spp_nutr_density.R")
# source ("Scripts/Plot_nutr_demand_met.R")
# source ("Scripts/Plot_n_inds_reqs_met.R")


# vector of country names
demonstration_countries <- c("Chile", "Peru", "Indonesia", "Sierra Leone")
```

## Nutrient density of major species

These cobweb plots tell us about the nutrient content of the most-caught species in each country. These are slightly different from a version I showed earlier, and I think more relevant. Now the y axis is the % recommended dietary allowance (RDA) for a child age 6mos to 8 years old from a 100g serving of that species. 100% means that eating a single serving of that species would satisfy a child's daily needs for that nutrient. 



```{r spider_plots, echo = FALSE}

radar_plots <- lapply (demonstration_countries, plot_spp_nutr_radar, n_spp = 5)
#grid.arrange (radar_plots[[1]], radar_plots[[2]],radar_plots[[3]],radar_plots[[4]], ncol =1)

radar_plots[[1]]

```
```{r, echo = FALSE}
# surely a better way of doing this??
radar_plots[[2]]
```
```{r, echo = FALSE}
radar_plots[[3]]
```

Note: Acetes japonicus and P. trituberculatus are both crustaceans and thus have identical nutrient profiles from GENuS data. Last graph had a different fish, was 5th most caught with just 2012 while P. trituberculatus was 6th. Can look into further..
```{r, echo = FALSE}
radar_plots[[4]]
```

### How much are the most nutritious species contributing to catch?

```{r catch_vs_dens}

```

