---
title: "Indonesia regional spp"
date: 2/1/2023
output: html_notebook
---

Indonesia team sent an updated list of species on Jan 24, 2023 for a presentation on 2/2/23

•	Kakap merah (red snapper) Lutjanus gibbus
•	Kerapu lumpur (Orange-spotted grouper) Epinephelus coioides
•	Ikan Kembung (Indian and short mackerel) Rastrelliger kanagurta and R. brachysoma : Regarding to this one, can we mention some other small pelagic species? Including Layang (Indian scad) Decapterus russeli and Teri (Anchovy) Stholephorus sp.
•	Kepiting bakau (mud crab) Schylla serrata
•	Tuna sirip kuning (yellowfin tuna) Thunnus albacares



```{r data_import}
library (tidyverse)
library (AFCD)
library (stringr)

# this will also bring in fishnutr data and RNI data
source ("Scripts/Function_convert_catch_amt_children_fed.R")

# priority species from regional teams
pri_spp <- read.csv("Data/regional_teams_priority_spp.csv") 
pri_spp_indo <- filter (pri_spp, country == "Indonesia")

# SAU landings data
sau_indo <- read.csv("Data/SAU EEZ indonesia.csv") 
sau_indo_2019 <- sau_indo %>% filter (year == 2019)

# nutricast/fishcast projections
ds_spp <- read_rds("Data/Free_etal_proj_smaller.Rds")
ds_spp_indo <- filter (ds_spp, country == "Indonesia")

# exports from ARTIS, emailed 1/28/23
exports_indo <- read_csv("Data/20230125_edf_ARTIS_indo.csv")
```

```{r data_avail}

fishnutr_mu %>%
  right_join (pri_spp, by = "species")  %>%
  pivot_longer (Selenium_mu:Vitamin_A_mu,
                names_to = "nutrient",
                values_to = "amount") %>%
  filter (is.na(amount)) %>% 
  select (country, comm_name, species) %>% 
  distinct()

fishnutr_mu %>% filter (grepl("Stolephorus", species))


sau_indo_2019 %>%
  filter (scientific_name %in% pri_spp_indo$species) %>% #pull (scientific_name) %>% unique()
  group_by (scientific_name) %>%
  summarise (tonnes = sum (tonnes, na.rm = TRUE))

sau_2019 %>% filter (grepl ("Stolephorus", scientific_name)) %>% View()
sau_2019 %>% filter (grepl ("Lutjanus", scientific_name)) %>% View()
sau_2019 %>% filter (grepl ("Epinephelus", scientific_name)) %>% View()
sau_indo_2019 %>% filter (grepl ("Epinephelus", scientific_name)) %>% group_by (scientific_name) %>% summarise (tonnes = sum (tonnes))

ds_spp_indo %>% 
  filter (species %in% pri_spp_indo$species) %>% pull (species) %>% unique()
ds_spp_indo %>% filter (grepl("Stolephorus", species)) %>% View()

```

FishNutrients has 6: doesn't have scylla (crab). 
8 Stolephorus options, all of which occur in Indonesia according to FishBase. Also Encrasicholina puncifer or E. japonicus
Just take the average of all of them! 

SAU: Has 5 spp: R. brach, R. kanagurat, D. russ, S. serrata, T. alb. 
But D. russelli; R. kanagurta is all foreign catch (thai); S. serrata very litte (42 tonnes)
SAU has general Stolephorus (calls it garment anchovy) and Lutjanus; no other Lutjanus species.
Also has general Epinephelus (only at genus level, no other species), but only Malaysian catch

Nutricast has 6: R. brach, R. kana, S. serrat, T. alb, D. russ, L. gib; NO: Stolephorus but does have Engraulis japonicus; Encrasicholina punctifer (these are the two with anchovy in common name); NO E. coides but 6 other Ephinephelus spp. Fishbase says coioides often misidentified as E. tauvina which is in ds

Exports has 5: R. brach, R. kana, S. serrat, Stolephorus, T. alb
*Might work if I ask for Lutjanus

```{r indo_stolep_possibilities}

# make a vector of possible stolep spp to pull from fishnutr

indo_stolephorus <- fishnutr_mu %>% 
  filter (grepl("Stolephorus", species) | species == "Encrasicholina punctifer" | species =="Engraulis japonicus") %>%
  pull (species) %>%
  unique ()

saveRDS (indo_stolephorus, file = "Data/indo_stolephorus.Rds")

```

```{r}
sau_indo_2019 %>%
    filter (scientific_name %in% pri_spp_indo$species | scientific_name == "Lutjanus" | scientific_name == "Stolephorus" | scientific_name == "Epinephelus") %>%
  ggplot (aes (x= scientific_name, y = tonnes)) +
  geom_col() +
  theme_bw()
    
```

write out nutrition facts for L. gibbus
```{r}
fishnutr_long %>%
  filter (species == "Lutjanus gibbus") %>%
  left_join(rni_child, by = "nutrient") %>%
  mutate (perc_rni = amount/RNI * 100)
```

