---
title: "Peru D. gigas case study"
author: "JGM"
date: "2023-04-03"
output: html_document
---

```{r setup, include=FALSE}
library (tidyverse)
library (stringr)

# function for converting catch in mt to children fed ----
# this will also bring in fishnutr data and RNI data
source ("Scripts/Function_convert_catch_amt_children_fed.R")

# SAU data ----

# as of 10/25/22 just 2019 data, suggested by Deng Palomares. Clipped in SAU_explore.R
sau_2019 <- readRDS("Data/SAU_2019.Rds")

# or mean of most recent 5 years *doesn't currently include indonesia* 
sau_2015_2019 <- readRDS("Data/SAU_2015_2019.Rds")


# exports ARTIS ----
# emailed 10 19 2022
exports <- read_csv ("Data/20221019_edf_ARTIS_snet.csv")

# take a five year mean of proportion exported
exports_5yr_mean <- exports %>%
  filter (between (year, 2015, 2019)) %>%
  group_by (exporter_iso3c, sciname) %>%
  summarise (mn_prop_exp = mean (exports_percent_of_prod, na.rm = TRUE)/100) %>%
  ungroup() %>%
  mutate (species = str_to_sentence(sciname),
          # just doing country names by hand...
          country = case_when (
            exporter_iso3c == "CHL" ~ "Chile",
            exporter_iso3c == "IDN" ~ "Indonesia",
            exporter_iso3c == "PER" ~ "Peru", 
            exporter_iso3c == "MEX" ~ "Mexico",
            exporter_iso3c == "SLE" ~ "Sierra Leone"
          ), .keep = "unused") 

# expressed as catch ratios relative to base year for midcentury and end century, can multiply by landings
catch_upside_relative <- readRDS("Data/nutricast_upside_relative.Rds")
```

# nutrient content
```{r print_nutrient_content}
compiled_nutr %>% 
  filter (species == "Dosidicus gigas") %>%
  left_join(rni_child, by = "nutrient") %>%
  mutate (perc_rni = amount/RNI * 100) %>%
  arrange (desc (perc_rni))
```
```{r peru_inadequate_intake}
# this is already in percent, don't multiply by 100
pii <- read_csv ("Data/Beal_2017_SI_inadequate_micronutrients.csv")

peru_pii <- pii %>% 
  filter (Country == "Peru", Year == 2011) %>%
  select (Fortification, Micronutrient, `Prevalence of Inadequate Intake`) 


peru_pii %>%
  arrange (desc (round(`Prevalence of Inadequate Intake`, 2)))


```
```{r compare_radar_barchart}
library (ggradar)


spp_rni_met <- compiled_nutr %>%
  filter (species %in% c("Dosidicus gigas", "Trachurus murphyi", "Merluccius gayi peruanus", "Coryphaena hippurus", "Oncorhynchus tshawytscha")) %>%
  # join to rni data
  left_join (rni_child, by = "nutrient") %>%
  
  # this would be the percentage of your daily requirement you could get from a 100g serving of each species. cap at 100%
  mutate (perc_rni = amount/RNI * 100,
          perc_rni = ifelse (perc_rni > 100, 100, perc_rni),
          nutrient = 
            case_when (nutrient == "Vitamin_A" ~ "Vit A",
                       nutrient == "Omega_3" ~ "Omg 3",
                       TRUE ~ nutrient)) %>%
  ungroup()


nutr_radar_plot <- spp_rni_met  %>%
  # filter (species %in% c("Engraulis ringens", "Trachurus murphyi", "Merluccius gayi gayi"), !nutrient %in% c("Protein", "Selenium")) %>%
   mutate(
      spp_short = ifelse (
        grepl(" ", species),
        paste0 (substr(species, 1, 1), ". ", str_split_fixed (species, " ", 2)[,2]),
        species)) %>%
  select (spp_short, nutrient, perc_rni) %>%
  pivot_wider (
    names_from = nutrient,
    values_from = perc_rni) 


#png ("Figures/Squid_radar.png", res = 300, width = 5, height = 5, units = "in")  
ggradar(nutr_radar_plot,
        grid.min = 0, grid.max = 100, 
        group.point.size = 2,
        group.line.width = 1,
        legend.text.size = 8,
        axis.label.size = 4,
        grid.label.size = 4,
        legend.position = "bottom") +
  
  theme (plot.title = element_text (size = 14))
#dev.off()

png ("Figures/Squid_nutr_dodge_bar_sm_salmon.png", width = 7, height = 4, units = "in", res = 300)
 spp_rni_met  %>%
   filter (!nutrient %in% c("Protein", "Selenium")) %>%
    group_by (species) %>%
   # samuel asked abt omega specifically, make a rank based on omg3?
    mutate (micronutrient_density = sum (perc_rni),
            #omg_ranking = 
            spp_short = ifelse (
              grepl(" ", species),
              paste0 (substr(species, 1, 1), ". ", str_split_fixed (species, " ", 2)[,2]),
              species)
    ) %>%
    ungroup() %>%
    
    ggplot (aes (x = reorder(spp_short, -micronutrient_density), fill = nutrient, y = perc_rni)) +
    geom_col (position = "dodge") +
    theme_bw() +
    labs (x = "", y = "% Child RNI met per 100g serving", fill = "Nutrient") +
    #ylim (c(0,100)) +
    ggtitle ("Nutrient content of selected species") +
    theme ( 
      axis.text.y = element_text (size = 13),
      axis.text.x = element_text (size = 11),
      axis.title = element_text (size = 16),
      strip.text = element_text(size = 16),
      legend.text = element_text (size = 12),
      legend.title = element_text (size = 14),
      plot.title = element_text (size = 18))
 dev.off()

```

samuel asked if we could just say in words, calculate amt, say has X% more than another species
```{r omg_percent_compare}

spp_omg <- spp_rni_met %>%
  filter (nutrient == "Omg 3")

#compiled_nutr %>% filter (species == "Engraulis ringens")

```

D. gigas has 0.6 g omega 3 per servings, 54.5% RNI. 

That's `r round(spp_omg$amount[which(spp_omg$species == "Dosidicus gigas")]/spp_omg$amount[which(spp_omg$species == "Oncorhynchus tshawytschas")], 2)`


About 2x as much as jurel, about 4x as much as hake, slightly more than mahi mahi. salmon has 2.5x as much. anchovy has 1.02, 1.7x as much

```{r catch rnis}

# just pota for Peru, 2019
  pota_catch <- sau_2019 %>% filter (country == "Peru", fishing_entity == "Peru", species == "Dosidicus gigas") %>%
    group_by (species) %>% summarise (tonnes = sum (tonnes))
  
  pota_2019 <- calc_children_fed_func(species = "Dosidicus gigas", amount = pota_catch$tonnes, country = "Peru")
  
  png ("Figures/Peru_SAU_2019_pota_children_fed.png", res = 300, width = 5.5, height = 5, units = "in")  
  pota_2019 %>%
   # filter (nutrient != "Protein") %>%
    ggplot (aes (x = nutrient, y = children_fed / 1000000, fill = nutrient)) +
    geom_col() +
    theme_bw() +
    labs (x = "", y = "") +
    theme (axis.text.x = element_blank (),
           axis.text.y = element_text (size = 15),
           legend.position = "none")
  
  dev.off()

```


```{r trade_levers}
z <- sau_2019 %>%
  filter (country == "Peru", species == "Dosidicus gigas") %>%
  left_join (exports_5yr_mean, by = c("species", "country")) %>%
  replace_na (list(mn_prop_exp = 0)) %>%
  group_by (species) %>%
  summarise (foreign_catch = sum (tonnes[fishing_entity != "Peru"]),
             domestic_catch = sum (tonnes[fishing_entity == "Peru"]) * (1-mn_prop_exp),
             exported = sum (tonnes[fishing_entity == "Peru"]) * mn_prop_exp) %>%
  # creating many repeats, not sure why
  # warning about reframe()
  distinct () %>%
  # pivot longer
  pivot_longer (foreign_catch:exported,
                names_to = "lever",
                values_to = "catch_mt") %>%
  mutate (children_fed = pmap (list (species = species, amount = catch_mt, country_name = "Peru"), calc_children_fed_func)) %>%
  unnest (cols = c(children_fed)) %>%
  rename (mt = catch_mt)

z$lever <- factor (z$lever, levels = c ("exported", "foreign_catch", "domestic_catch"))

Selenium <- FALSE
if (Selenium == TRUE) {omit_nutrients <- "Protein"} else {omit_nutrients <- c("Protein", "Selenium")}


#png ("Figures/Peru_trade_levers_squid.png", width = 5, height = 4, unit = "in", res = 300)
 z %>%
  filter (!nutrient %in% omit_nutrients) %>%
  group_by (nutrient, lever) %>%
  summarise (children_fed = sum (children_fed, na.rm = TRUE)) %>%
  ggplot (aes (y = children_fed/1000000, x = reorder(nutrient, -children_fed), fill = lever)) +
  geom_col () +
  theme_bw()+
  scale_fill_grey(start = 0.8, end = 0.2) +
  labs (y = "Child RNIs met, millions", x = "", fill = "Policy lever") +
  ggtitle ("Allocative losses, trade/foreign catch") +
  theme ( 
    axis.text.y = element_text (size = 11),
    axis.text.x = element_text (size = 11),
    axis.title = element_text (size = 16),
    legend.text = element_text (size = 11),
    legend.title = element_text (size = 13),
    plot.title = element_text (size = 18),
    legend.position = c(0.8, 0.8))
 #dev.off()
 
 # print percentages
exports_5yr_mean %>% filter (country == "Peru", species == "Dosidicus gigas")

b <- sau_2019 %>%
  filter (country == "Peru", species == "Dosidicus gigas") %>%
  left_join (exports_5yr_mean, by = c("species", "country")) %>%
  replace_na (list(mn_prop_exp = 0)) %>%
  group_by (species) %>%
  summarise (foreign_catch = sum (tonnes[fishing_entity != "Peru"]),
             domestic_catch = sum (tonnes[fishing_entity == "Peru"]) * (1-mn_prop_exp),
             exported = sum (tonnes[fishing_entity == "Peru"]) * mn_prop_exp) %>%
  # creating many repeats, not sure why
  # warning about reframe()
  distinct ()
 

# which fishing nations?
sau_2019 %>%
    filter (country == "Peru", species == "Dosidicus gigas") %>%
  group_by(fishing_entity) %>%
  summarise (tot = sum (tonnes))
```

```{r end_use_levers}

c <- sau_2019 %>%
   filter (country == "Peru", species == "Dosidicus gigas") %>%
 
   group_by (species, end_use_type) %>%
   summarise (catch_mt = sum (tonnes, na.rm = TRUE)) %>%

   mutate (children_fed = pmap (list (species = species, amount = catch_mt, country_name = "Peru"), calc_children_fed_func)) %>%
   unnest (cols = c(children_fed)) %>%
   rename (mt = catch_mt)
 
 c$end_use_type <- factor (c$end_use_type, levels = c ("Fishmeal and fish oil", "Other", "Direct human consumption"))
 
 #if (Selenium == TRUE) {omit_nutrients <- "Protein"} else {omit_nutrients <- c("Protein", "Selenium")}
 png ("Figures/Peru_enduse_levers_squid.png", width = 5, height = 4, unit = "in", res = 300)
 c %>%
   filter (!nutrient %in% omit_nutrients, !is.na (end_use_type)) %>%
   group_by (nutrient, end_use_type) %>%
   summarise (children_fed = sum (children_fed, na.rm = TRUE)) %>%
   ggplot (aes (x = reorder(nutrient, -children_fed), y = children_fed/1000000, fill = end_use_type)) +
   geom_col () +
   theme_bw()+
   scale_fill_grey(start = 0.8, end = 0.2) +
   labs (y = "Child RNIs met, millions", x = "", fill = "Policy lever") +
   ggtitle ("Allocative losses, end uses") +
   theme ( 
     axis.text.y = element_text (size = 11),
     axis.text.x = element_text (size = 11),
     axis.title = element_text (size = 16),
     legend.text = element_text (size = 11),
     legend.title = element_text (size = 13),
     plot.title = element_text (size = 18),
     legend.position = c(0.8, 0.8))
 
dev.off()

c %>% filter (nutrient == "Zinc")
c %>% filter (nutrient == "Omega_3")

```
```{r end_use_by_country}

d <- sau_2019 %>%
   filter (country == "Peru", species == "Dosidicus gigas") %>%
 
   group_by (species, end_use_type, fishing_entity, fishing_sector) %>%
   summarise (catch_mt = sum (tonnes, na.rm = TRUE))  

```

Samuel feedback 4/11/23: no peruvian industrial fleets for Pota. try looking at past years and reconstruction method. 

In De la puente 2020 reconstruction--d. gigas in industrial fleets is bycatch in domestic fleet. Assumed that the catch rate would be 2% of hake landings in trawl nets and 1% of chub/jack macerel purse seines, 0.01% of anchovy/sardine landings. assumed all industrial catch was fishmeal fishoil bc bycatch. 

assume some use as bait for mahi mahi, blue shark, mako in longlines, but not sure how this is factored in to landings?

importantly--they account for eyes, ink sack, and guts being discarded at sea. assume 15% by weight, so assume only 85% of total catch was reported in landings. but I think this means they padded the reported landings, not that I need to account for this with edible weight. 
```{r examine_sau_pattners}
dg_sau_ts <- sau_2015_2019 %>% filter (species == "Dosidicus gigas", country== "Peru")

dg_sau_ts %>%
  filter (!fishing_sector == "Subsistence") %>%
  ggplot (aes (x = year, y = tonnes, fill = end_use_type)) +
  geom_col() +
  theme_bw()+
  facet_wrap (~fishing_sector) +
  ggtitle ("SAU D. gigas landings, Peru, 2019 reconstruction\ndownloaded 10/22")


```

so this is very weird, saying that most of the artisanal is fishmeal, and most of the industrial is DHC. 

look at older dataset, before 2019 reconstruction
```{r older_sau}

download_older_full <- read.csv ("Data/SAU_EEZ_landings.csv")
dg_sau_older <- download_older_full %>% filter (scientific_name == "Dosidicus gigas", area_name== "Peru")

png("Figures/SAU_Peru_Dgigas_sector_2018.png", width = 7, height =5, unit = "in", res = 300)
dg_sau_older %>%
  filter (!fishing_sector == "Subsistence", year > 2000) %>%
  ggplot (aes (x = year, y = tonnes, fill = end_use_type)) +
  geom_col() +
  theme_bw()+
  facet_wrap (~fishing_sector) +
  ggtitle ("SAU D. gigas landings, Peru, 2018 reconstruction\ndownloaded 03/22")
dev.off()

```
also look at full years for 2019 reconstruction
```{r 2019_reconstruction_longer_ts}
download_2019_full <- read.csv("Data/SAU EEZ 2019.csv") 

dg_sau_2019_full <- download_2019_full %>% filter (scientific_name == "Dosidicus gigas", area_name== "Peru")

png("Figures/SAU_Peru_Dgigas_sector_2019.png", width = 7, height =5, unit = "in", res = 300)
dg_sau_2019_full %>%
  filter (!fishing_sector == "Subsistence", year > 2000) %>%
  ggplot (aes (x = year, y = tonnes, fill = end_use_type)) +
  geom_col() +
  theme_bw()+
  facet_wrap (~fishing_sector) +
  ggtitle ("SAU D. gigas landings, Peru, 2019 reconstruction\ndownloaded 10/22")
dev.off()

# what about foreign
png("Figures/SAU_Peru_Dgigas_end_use_foreign_ts.png", width = 5, height =5, unit = "in", res = 300)
dg_sau_2019_full %>%
  filter (!fishing_sector == "Subsistence", year > 2009, !fishing_entity == "Peru") %>%
  ggplot (aes (x = year, y = tonnes/1000000, fill = end_use_type)) +
  geom_col() +
  theme_bw()+
  facet_wrap (~fishing_sector) +
  scale_x_continuous(breaks = c (2010, 2014, 2018)) +
  labs (y = "million tonnes") +
  ggtitle ("SAU D. gigas foreign landings")
dev.off()

png("Figures/SAU_Peru_Dgigas_end_use_domestic_ts.png", width = 7, height =5, unit = "in", res = 300)

dg_sau_2019_full %>%
  filter (!fishing_sector == "Subsistence", year > 2009, fishing_entity == "Peru") %>%
  ggplot (aes (x = year, y = tonnes/1000000, fill = end_use_type)) +
  geom_col() +
  theme_bw()+
   labs (y = "million tonnes") +
  scale_x_continuous(breaks = c (2010, 2014, 2018)) +
  facet_wrap (~fishing_sector) +
  ggtitle ("SAU D. gigas domestic landings")
dev.off()

 
```
okay. so the peru industrial is tiny, and mostly FMFO. the industrial FMFO is foreign catch, which I guess is DHC.

```{r nutricast_ts}
ds_spp <- readRDS("Data/Free_etal_proj_smaller.Rds")


dat <- ds_spp%>% 
    filter (country == "Peru", species == "Dosidicus gigas", year  >2030)
  
  dat$scenario <- factor (dat$scenario, levels = c ("No Adaptation", "Productivity Only", "Full Adaptation"))
  
  p <- dat %>% ggplot (aes (x = year, y = catch_mt/1000, col = scenario)) +
    geom_line() +
    theme_bw() +
    facet_grid (~ rcp, scales = "free_y") +
    labs (x = "", y = "Catch, metric tons (1000s)", col = "Mgmt scenario") +
    ggtitle ("Free et al. (2020) projections for D. gigas") +
    theme (plot.title = element_text (size = 14),
           axis.text = element_text (size = 12))

  png ("Figures/Peru_squid_nutricast_ts.png", res = 300, width = 12, height = 5, units = "in")  
  print (p) +
    theme(strip.text = element_text (size = 12),
          legend.title = element_text (size = 13),
          legend.text = element_text (size = 11),
          axis.title = element_text (size = 14))
  dev.off()
  
  
  
# samuel asked for the full time series  
  
  dat <- ds_spp%>% 
    filter (country == "Peru", species == "Dosidicus gigas")
  
  dat$scenario <- factor (dat$scenario, levels = c ("No Adaptation", "Productivity Only", "Full Adaptation"))
  
  p <- dat %>% ggplot (aes (x = year, y = catch_mt/1000, col = scenario)) +
    geom_line() +
    theme_bw() +
    facet_grid (~ rcp, scales = "free_y") +
    labs (x = "", y = "Catch, metric tons (1000s)", col = "Mgmt scenario") +
    ggtitle ("Free et al. (2020) projections for D. gigas") +
    theme (plot.title = element_text (size = 14),
           axis.text = element_text (size = 12))

  png ("Figures/Peru_squid_nutricast_ts_2020.png", res = 300, width = 12, height = 5, units = "in")  
  print (p) +
    theme(strip.text = element_text (size = 12),
          legend.title = element_text (size = 13),
          legend.text = element_text (size = 11),
          axis.title = element_text (size = 14))
  dev.off()
  
  # also zoom in on midcentury
  
  png ("Figures/Peru_squid_nutricast_ts_midcentury.png", res = 300, width = 12, height = 5, units = "in")  
  dat %>% 
    filter (year %in% 2051:2060) %>%
    ggplot (aes (x = year, y = catch_mt/1000, col = scenario)) +
    geom_line() +
    theme_bw() +
    facet_grid (~ rcp, scales = "free_y") +
    labs (x = "", y = "Catch, metric tons (1000s)", col = "Mgmt scenario") +
    ggtitle ("Free et al. (2020) projections for D. gigas, mid-century") +
    theme (plot.title = element_text (size = 14),
           axis.text = element_text (size = 12))
  dev.off()
  
   png ("Figures/Peru_squid_nutricast_ts_endcentury.png", res = 300, width = 12, height = 5, units = "in")  
  dat %>% 
    filter (year %in% 2091:2100) %>%
    ggplot (aes (x = year, y = catch_mt/1000, col = scenario)) +
    geom_line() +
    theme_bw() +
    facet_grid (~ rcp, scales = "free_y") +
    labs (x = "", y = "Catch, metric tons (1000s)", col = "Mgmt scenario") +
    ggtitle ("Free et al. (2020) projections for D. gigas, end-century") +
    theme (plot.title = element_text (size = 14),
           axis.text = element_text (size = 12))
  dev.off()
  
```
```{r nutricast profit ts}
ds_squid_long <- ds_spp%>% 
    filter (country == "Peru", species == "Dosidicus gigas") %>%
  pivot_longer(range_km2:ffmsy, 
               names_to = "variable",
               values_to = "value")
  
  ds_squid_long$scenario <- factor (ds_squid_long$scenario, levels = c ("No Adaptation", "Productivity Only", "Full Adaptation"))
  
  ds_squid_long$variable <- factor (ds_squid_long$variable, levels = c ("biomass_mt", "catch_mt", "profits_usd", "range_km2", "msy_mt", "bbmsy", "ffmsy"))

  png ("Figures/Peru_squid_nutricast_ts_2020_allvar.png", res = 300, width = 12, height = 6, units = "in") 
  p <- ds_squid_long %>% 
    filter (variable %in% c ("catch_mt", "biomass_mt", "profits_usd")) %>%
    ggplot (aes (x = year, y = value/1000000, col = scenario)) +
    geom_line () +
    facet_grid (variable ~ rcp, scales = "free_y") +
    theme_bw() +
    labs (x = "", y = "") +
    ggtitle (paste0 ("Free et al. (2020) projections, all variables (millions)")) +
    theme (plot.title = element_text (size = 14),
           axis.text = element_text (size = 12))
  print(p)
  dev.off()

```

```{r nutrient_upside}
squid_upside_nutr <- sau_2019 %>%
  filter(country == "Peru", species == "Dosidicus gigas") %>%
  group_by (country, species) %>%
  summarise (total_tonnes = sum (tonnes)) %>%
  ungroup() %>%
  left_join (catch_upside_relative, by = c("country", "species")) %>%
  mutate (# multiply ratio by current landings
    across(bau_ratio_midcentury:adapt_ratio_endcentury, ~.x * total_tonnes),
    #convert to upside, subtract 
    mey_2050 = mey_ratio_midcentury - bau_ratio_midcentury,
    mey_2100 = mey_ratio_endcentury - bau_ratio_endcentury,
    adapt_2050 = adapt_ratio_midcentury - bau_ratio_midcentury,
    adapt_2100 = adapt_ratio_endcentury - bau_ratio_endcentury) %>%
  
  select (country, rcp, species, mey_2050:adapt_2100) %>%
  pivot_longer(mey_2050:adapt_2100, 
               names_to = c("scenario", "period"),
               names_sep = "_",
               values_to = "tonnes")  %>%
    # get rid of non-matching species, NAs
    filter (!is.na (rcp)) %>%
    # convert to nutrients
    mutate (children_fed = pmap (list (species = species, amount = tonnes, country_name = country), calc_children_fed_func)) %>%
    unnest(cols = c(children_fed),  names_repair = "check_unique") 
  
  
  # fix ratios
squid_upside_nutr$scenario <- factor(squid_upside_nutr$scenario, levels = c ("mey", "adapt"))

png ("Figures/Peru_nutricast_upside_squid_allRCPs_midcentury.png", width = 13, height = 7, units= "in", res = 300)
squid_upside_nutr %>%
  filter (period == "2050") %>%
    # preliminary plot
    ggplot (aes (x = reorder(nutrient, -children_fed), y = children_fed/1000000, fill = scenario)) +
    geom_col (position = "dodge") +
    geom_hline (yintercept = 0, lty = 2) +
    facet_wrap (~rcp) +
    theme_bw() +
    # roughly match colors from gaines et al
    scale_fill_manual (values = c ("mediumseagreen", "dodgerblue4"), labels = c("Productivity Only", "Full Adaptation")) +
    labs (y = "Change in # children fed, millions", x = "", fill = "Management \nstrategy") +
    ggtitle ("Nutrition upside from climate-adaptive management, midcentury") +
    theme (strip.text = element_text (size = 12),
          legend.title = element_text (size = 13),
          legend.text = element_text (size = 11),
          axis.title = element_text (size = 14),
         axis.text = element_text (size = 12),
         plot.title = element_text (size = 16))
  
dev.off()

png ("Figures/Peru_nutricast_upside_squid_allRCPs_endcentury.png", width = 13, height = 7, units= "in", res = 300)
squid_upside_nutr %>%
  filter (period == "2100") %>%
    # preliminary plot
    ggplot (aes (x = reorder(nutrient, -children_fed), y = children_fed/1000000, fill = scenario)) +
    geom_col (position = "dodge") +
    geom_hline (yintercept = 0, lty = 2) +
    facet_wrap (~rcp) +
    theme_bw() +
    # roughly match colors from gaines et al
    scale_fill_manual (values = c ("mediumseagreen", "dodgerblue4"), labels = c("Productivity Only", "Full Adaptation")) +
    labs (y = "Change in # children fed, millions", x = "", fill = "Management \nstrategy") +
    ggtitle ("Nutrition upside from climate-adaptive management, 2100") +
  theme (strip.text = element_text (size = 12),
          legend.title = element_text (size = 13),
          legend.text = element_text (size = 11),
          axis.title = element_text (size = 14),
         axis.text = element_text (size = 12),
         plot.title = element_text (size = 16))
  
dev.off()



#display value
squid_upside_nutr %>% filter (nutrient == "Zinc")

# display percent yield?
pri_spp_catch_upside <- ds_spp %>% 
  filter(country == "Peru", species == "Dosidicus gigas") %>%
  mutate (
    period = case_when (
      year %in% c(2026:2035) ~ "2026-2035",
      year %in% c(2051:2060) ~ "2051-2060",
      year %in% c(2091:2100) ~ "2091-2100"
    )) %>%
  filter (!is.na (catch_mt), !is.na (period), scenario %in% c("No Adaptation", "Productivity Only", "Full Adaptation")) %>%
  #take mean projected catch for the decade period
  group_by (country, rcp, period, species, scenario) %>%
  summarise (catch_mt = mean (catch_mt)) %>%
  ungroup() %>%
  # find difference among scenarios--absolute and percent diff
  group_by (country, rcp, period, species) %>%
  summarize (mey_diff_mt = catch_mt[scenario == "Productivity Only"] - catch_mt[scenario == "No Adaptation"],
             mey_diff_percent = (catch_mt[scenario == "Productivity Only"] - catch_mt[scenario == "No Adaptation"])/catch_mt[scenario == "No Adaptation"] * 100,
             adapt_diff_mt = catch_mt[scenario == "Full Adaptation"] - catch_mt[scenario == "No Adaptation"],
             adapt_diff_percent = (catch_mt[scenario == "Full Adaptation"] - catch_mt[scenario == "No Adaptation"])/catch_mt[scenario == "No Adaptation"] * 100) %>%
  ungroup()


pri_spp_catch_upside %>% filter (period == "2051-2060")
```

