---
title: "SAU data exploration"
date: 4/5/22
output: html_notebook
---

SAU data downloaded 3/31/22

Columns:
area_name - country name
area_type - eez
data_layer "Reconstructed domestic catch" "Inferred foreign catch"       "Assigned Tuna RFMO catch" 
  Maybe limit to domestic catch? But could compare nutrition yield of domestic and foreign?
uncertainty_score - 1-4
year 1950-2018, yearly. probably use 2000-2018
scientific_name - includes family/group names, not identified. separated with space
common_name
functional_group
commercial_group
fishing_entity - country names, maybe flag ship
fishing_sector - "Artisanal"    "Subsistence"  "Recreational" "Industrial" --> take out rec
catch_type - landings / discards --> take out discards
reporting_status - reported, unreported keep both
gear_type
end_use_type - "Direct human consumption" "Fishmeal and fish oil"    "Other"                    "Discards" --> only DHC
tonnes
landed_value



```{r setup, echo = FALSE}
knitr::opts_chunk$set(echo = FALSE)

library (tidyverse)

SAU <- read_csv ("Data/SAU_EEZ_landings.csv")
```
start with latin america

# Chile

chile mainland only


```{r chile_ds}
sau_chl <- SAU %>%
   filter (between (year, 2000, 2015), 
          tonnes > 0, grepl ("Chile", area_name))

unique (sau_chl$fishing_entity) # lots of different ones
sau_chl_dom <- sau_chl %>% filter (data_layer == "Reconstructed domestic catch")
unique (sau_chl_dom$fishing_entity) # just chile

# end use and landings by sector
sau_chl %>%
  ggplot (aes (x = year, y = tonnes, fill = end_use_type)) +
  geom_bar(stat = "identity") +
  facet_wrap (~fishing_sector) +
  theme_bw()

# landings vs discards
sau_chl %>%
  ggplot (aes (x = year, y = tonnes, fill = catch_type)) +
  geom_bar(stat = "identity") +
  facet_wrap (~fishing_sector) +
  theme_bw()

# reported, unreported
sau_chl %>%
  ggplot (aes (x = year, y = tonnes, fill = reporting_status)) +
  geom_bar(stat = "identity") +
  facet_wrap (~fishing_sector) +
  theme_bw()

# domestic, foreign
  sau_chl %>%
  ggplot (aes (x = year, y = tonnes, fill = data_layer)) +
  geom_bar(stat = "identity") +
  facet_wrap (~fishing_sector) +
  theme_bw()
  
  # weird anchovy situation
  sau_chl %>%
    filter (scientific_name == "Engraulis ringens") %>%
  ggplot (aes (x = year, y = tonnes, fill = data_layer)) +
  geom_bar(stat = "identity") +
  facet_wrap (~fishing_sector) +
  theme_bw()
```
Industrial is almost all fishmeal. big declining trend, and very little catch 2016-2018, nearly all DHC. big drop 2015, and then no fishmeal 2016-2018. Artisanal is growing and mostly DHC.  

Very few discards, limited unreported. little foreign

anchovy in artisanal only starting in 2015, concurrent with a big decline in industrial. looks like general decline in 2015. [but this was also el nino?]

### major species by sector

https://www.subpesca.cl/portal/616/w3-propertyvalue-618.html

erica and sergio: Chilean Hake (Merluccius gayi), Austral Hake (Merluccius australis), and Conger Eel (congrio)
Genypterus blacodes
Genypterus maculatus

reineta: Brama australis
merluza: Merluccius gayi gayi
Macruronus magellanicus
Micromesistius australis
Merluccius australis

```{r chl_spp_sector}

sau_chl %>% filter (fishing_sector == "Artisanal") %>%
  group_by (scientific_name) %>%
  summarise (tot_cat = sum (tonnes)) %>% arrange (desc (tot_cat)) %>% View()
  slice_max (tot_cat, n = 10)

sau_chl %>% filter (fishing_sector == "Industrial") %>%
  group_by (scientific_name) %>%
  summarise (tot_cat = sum (tonnes)) %>%
  slice_max (tot_cat, n = 10)

# is gayi gai in artsanal??
sau_chl_ssf <- sau_chl %>% filter (fishing_sector == "Artisanal")
"Merluccius gayi gayi" %in% sau_chl_ssf$scientific_name # true, 21 most common

```
Artisanal catching giatn squid, n. crockeri (camotillo and bacaladillo?), anchov, unidentified. 
industrial anchovy, t. murphiyi, clupid, scomber japoin, sardonops, m gayi gayi, gient squid


Is there a clear breakdown in species caught in the different sectors?
```{r}
sau_chl %>%
  group_by (scientific_name) %>%
  summarise (prop_ind = sum (tonnes[fishing_sector == "Industrial"])/sum (tonnes),
             prop_ssf = sum (tonnes[fishing_sector == "Artisanal"]) / sum (tonnes)
  ) %>% arrange (desc (prop_ssf)) %>% View()
```

pretty strong distinction where there are a ton of species just in artisanal, 68, (76 > 99%) then a few 80s, 2 60s, one low 50 (Sprattus fuegensis). Prionace glauca 35% ssf, M gayi gayi 22%, Heterocarpus reedi 11%, then the rest 1-0%. 


#### major species caught in artisanal vs industrial sectors
```{r}
sum_artisanal <- sum (sau_chl$tonnes[which(sau_chl$fishing_sector == "Artisanal")])
sum_ind <- sum (sau_chl$tonnes[which(sau_chl$fishing_sector == "Industrial")])


chl_ssf_props <- sau_chl %>%
  filter (fishing_sector == "Artisanal") %>%
  group_by (scientific_name) %>%
  summarise (prop = round(sum(tonnes)/sum_artisanal * 100, 2),
             functional_group = first (functional_group)) %>%
  arrange (desc(prop))

chl_ind_props <- sau_chl %>%
  filter (fishing_sector == "Industrial") %>%
  group_by (scientific_name) %>%
  summarise (prop = round(sum(tonnes)/sum_ind * 100, 2),
            functional_group = first (functional_group)) %>%
  arrange (desc(prop))
```

#### which species do we lose from nutrition data
```{r}
chl_ind_props$scientific_name[which (!chl_ind_props$scientific_name %in% fishnutr_mu$species)]

chl_ssf_props %>% filter (!scientific_name %in% fishnutr_mu$species) %>% select (scientific_name, functional_group)
```
Industrial species--all families or inverts or unidentified. Could we replace "clupeidae" with sardinops sagax or other sardine? Strangomera bentincki, Sprattus fuegensis. S. sagax is definitely the most nutritious. Chose S. bentincki, medium? not in data already? S. bentincki is bigger than clup in ssf, not in industrial. 

```{r}
fishnutr_mu %>% filter (species %in% c("Sardinops sagax", "Sprattus fuegensis", "Strangomera bentincki"))

spp_micronutr_density %>% filter (species %in% c("Sardinops sagax", "Sprattus fuegensis", "Strangomera bentincki"))
```


Not missing any fish. all invertebrates and things not identified to species name. clup is the biggest priority to figure out

So, the problem is not the nutrition data. the problem is the Free data.

#### check join to free data
```{r  chl_free_check}

# projected nutritrient yield under management scenarios ----
# from Free data. Convert_catch_to_nutrients.R
ds_catch_nutr_yield_projected <- readRDS("Data/ds_catch_nutr_yield_projected.Rds")

ds_spp_names <- ds_catch_nutr_yield_projected %>% select (species) %>% distinct()


chl_ssf_props %>% filter (!scientific_name %in% ds_spp_names$species) %>% View()
chl_ssf_props %>% filter (!scientific_name %in% ds_spp_names$species) %>% summarise (n = sum(prop))
chl_ind_props %>% filter (!scientific_name %in% free_chl$species) %>% View()
chl_ind_props %>% filter (!scientific_name %in% free_chl$species) %>% summarise (n = sum(prop))
```
chl SSF: 64 species. marine unidentified 11% S. bentincki (clup) 3%,l. albus 8%, clup, brama australis 3%, dory gahi 3%, m. gayi gayi 1%. total 44.7%
ind mostly clup and m gayi gayi, 18%

Brama australis--> Brama brama? Atlantic pomfret. different resilience and temperature...
Merluccius gayi---M. australis?

Loxechinus albus is in the fishcast webtool though...as is Strangomera bentincki. L. albus in original ds_spp, not S. bentincki?
Lose urchin because no GENuS data for echinoderms. Don't have all the nutrients in AFCD yet. 

# Peru

```{r peru_ds}
sau_peru <- SAU %>%
   filter (between (year, 2000, 2015), 
          tonnes > 0, 
          area_name == "Peru", 
          fishing_sector %in% c ("Industrial", "Artisanal"))

unique (sau_peru$fishing_entity) # lots of different ones
sau_peru_dom <- sau_peru %>% filter (data_layer == "Reconstructed domestic catch")
unique (sau_peru_dom$fishing_entity) # just chile

unique (sau_peru$fishing_sector) # all 4 there, but negligbile sub/rec. especially rec

sau_peru %>%
  group_by (fishing_sector) %>%
  summarise (tot_cat = sum(tonnes))

# end use and landings by sector
sau_peru %>%
  filter (fishing_sector %in% c ("Industrial", "Artisanal")) %>%
  ggplot (aes (x = year, y = tonnes, fill = end_use_type)) +
  geom_bar(stat = "identity") +
  facet_wrap (~fishing_sector) +
  theme_bw()

# landings vs discards
sau_peru %>%
   filter (fishing_sector %in% c ("Industrial", "Artisanal")) %>%
  ggplot (aes (x = year, y = tonnes, fill = catch_type)) +
  geom_bar(stat = "identity") +
  facet_wrap (~fishing_sector) +
  theme_bw()

# reported, unreported
sau_peru %>%
   filter (fishing_sector %in% c ("Industrial", "Artisanal")) %>%
  ggplot (aes (x = year, y = tonnes, fill = reporting_status)) +
  geom_bar(stat = "identity") +
  facet_wrap (~fishing_sector) +
  theme_bw()

# domestic, foreign
  sau_peru %>%
     filter (fishing_sector %in% c ("Industrial", "Artisanal")) %>%
  ggplot (aes (x = year, y = tonnes, fill = data_layer)) +
  geom_bar(stat = "identity") +
  facet_wrap (~fishing_sector) +
  theme_bw()
```

similar declining trend of industrial, no fishmeal/oil in 2016-2018


```{r}
sau_peru %>%
group_by (scientific_name) %>%
  summarise (prop_ind = sum (tonnes[fishing_sector == "Industrial"])/sum (tonnes),
             prop_ssf = sum (tonnes[fishing_sector == "Artisanal"]) / sum (tonnes)
  ) %>% arrange (desc (prop_ssf)) %>% View()
```
not quite as stark as chile but a steady decline to 75% artisanal, then unidentified marine fish 52%, batoidea 47%, 	
Paralabrax humeralis 38%, Xiphias gladius 31%, Merluccius gayi peruanus 17% Scomber japonicus 13%, the rest under 10

#### species lost in free data
```{r}
sum_artisanal <- sum (sau_peru$tonnes[which(sau_peru$fishing_sector == "Artisanal")])
sum_ind <- sum (sau_peru$tonnes[which(sau_peru$fishing_sector == "Industrial")])


peru_ssf_props <- sau_peru %>%
  filter (fishing_sector == "Artisanal") %>%
  group_by (scientific_name) %>%
  summarise (prop = round(sum(tonnes)/sum_artisanal * 100, 2),
             functional_group = first (functional_group)) %>%
  arrange (desc(prop))

peru_ind_props <- sau_peru %>%
  filter (fishing_sector == "Industrial") %>%
  group_by (scientific_name) %>%
  summarise (prop = round(sum(tonnes)/sum_ind * 100, 2),
            functional_group = first (functional_group)) %>%
  arrange (desc(prop))
```

which are missing from free data
```{r}
peru_ssf_props %>% filter (!scientific_name %in% ds_spp_names$species) %>% View()
peru_ssf_props %>% filter (!scientific_name %in% ds_spp_names$species) %>% summarise (n = sum(prop))
peru_ind_props %>% filter (!scientific_name %in% ds_spp_names$species) %>% View()
peru_ind_props %>% filter (!scientific_name %in% ds_spp_names$species) %>% summarise (n = sum(prop))


```

Argopecten purpuratus 3%, makes sense, sea urchin. But Sarda chiliensis!!? 2%. m gayi peruanus. 18%, mostly unidentified
industrial, only 3.6%. 2% is clup


Sarda chiliensis do have nutrient info. In fishcast. So is M. gayi
have Sarda orientalis, and sarda (atlantic)



# sierra leone

for 2016-2018 compare
```{r sl_ds}

SAU %>%
  filter (year > 2000, 
          tonnes > 0, 
          area_name == "Sierra Leone", 
          fishing_sector %in% c ("Industrial", "Artisanal")) %>%
  ggplot (aes (x = year, y = tonnes, fill = end_use_type)) +
  geom_bar(stat = "identity") +
  facet_wrap (~fishing_sector) +
  theme_bw()


SAU %>%
  filter (year > 2000, 
          tonnes > 0, 
          grepl ("Indo", area_name), 
          fishing_sector %in% c ("Industrial", "Artisanal")) %>%
  ggplot (aes (x = year, y = tonnes, fill = end_use_type)) +
  geom_bar(stat = "identity") +
  facet_wrap (~fishing_sector) +
  theme_bw()
```
almost all DHC, more artisanal than industrial!

indo also 2016-2018 fishmeal

```{r}

```

